---
title: "Dataviz étranger en Suisse"
author: "Duc-Quang Nguyen"
date: "29 Aug 2016"
output: html_document
---

# Related

* [Press release](http://www.bfs.admin.ch/bfs/portal/fr/index/themen/01/01/new/nip_detail.Document.205535.pdf)

# Data 

* old 2014 [Population résidante permanente et non permanente selon les niveaux géographiques institutionnels, le sexe, la nationalité et l'âge](https://www.pxweb.bfs.admin.ch/Selection.aspx?px_language=fr&px_db=px-x-0103010000_101&px_tableid=px-x-0103010000_101\px-x-0103010000_101.px&px_type=PX)

* 2015 [Population résidante permanente et non permanente selon le canton, le sexe, l'autorisation de résidence, la classe d'âge et la nationalité](https://www.pxweb.bfs.admin.ch/DownloadFile.aspx?file=px-x-0103010000_101)

```{r settings}
loadPXdata <- F
data.file <- if(loadPXdata) "data/px-x-0103010000_101.px" else "input/0103010000_101_slim.csv"

n.threshold.merge <- 1000

library(pxR)
library(magrittr)
library(dplyr)
library(tidyr)
library(readr)
library(swiMap)
library(countrycode)

library(pipeR)
library(htmltools)
library(swiTheme)
library(swiRcharts)
library(sunburstR)
library(htmlwidgets)

```


```{r load data}
# load translations
txt <- read.csv("input/Sunburst CH foreigners 2015 - Sheet1.csv", row.names = 1, stringsAsFactors = F)

# discard incomplete translations
cidx <- unique(which(txt =="" | is.na(txt), T)[,2])
if(length(cidx > 0)) {
  warning(paste(colnames(txt)[cidx], collapse = "\t"), " languages will be discarded!", "\n")
  txt <- txt[,-cidx, drop = F]
}
colnames(txt)

# load data
if(loadPXdata) {
  px.read <- read.px("data/px-x-0103010000_101.px")
  data.read <- as.data.frame(px.read)
  #fr.metadata <- px.read[['VALUES.fr.']]
  en.metadata <- px.read[['VALUES.en.']]
  
  # filter data for the latest year, total age, ... keep only nationality and value
  data.read$Jahr <- as.numeric(as.character(data.read$Jahr))
  
  data.read %<>% filter(Jahr ==  max(data.read$Jahr)) %>%
    filter(Altersklasse ==  "Altersklasse - Total") %>%
    filter(Anwesenheitsbewilligung == "Anwesenheitsbewilligung - Total") %>%
    filter(Geschlecht == "Geschlecht - Total") %>%
    filter(Bevölkerungstyp == "Ständige Wohnbevölkerung") %>%
    filter(Kanton == "Schweiz") %>%
    rename(nationality = `Staatsangehörigkeit..Land.`) %>%
    select(nationality, value)
    
  # get the English nationality 
  nationality <- gsub('\"', '', unlist(strsplit(en.metadata$`Citizenship..Country.`, ",")))
  # renomve leading blank space
  nationality <- gsub("^ ?", "", nationality)
  names(nationality) <- levels(data.read$nationality)
  # convert nationlity to English
  data.read$nationality <- nationality[match(data.read$nationality, names(nationality))]
  
  # check total and remove total value
  check.tot <- data.read %>% filter(nationality != 'Citizenship - Total') %>%
    summarise(sum(value)) %>% unlist()
  check.tot2 <- data.read %>% filter(nationality == 'Citizenship - Total') %>% unlist()
  stopifnot(as.numeric(check.tot2[2]) == check.tot)
  
  data <- data.read %>% filter(nationality != 'Citizenship - Total')
  write.csv(data, file = "input/0103010000_101_slim.csv", row.names = F)
} else {
  data <- read.csv("input/0103010000_101_slim.csv", stringsAsFactors = F)
}
```

```{r wrange & structure data}

# Remove Suisses!
data <- filter(data, data$nationality != "Switzerland")
totForeigners <- sum(data$value)

## replace unknow origin by "Inconu"
data <- data %<>% 
  mutate(nationality = replace(nationality,
    nationality %in% c('Stateless', "Not attributable according to current borders", "No indication"),
    "unknown")) %>% 
  group_by(nationality) %>% summarise(value = sum(value, na.rm = T)) %>% ungroup()
stopifnot(sum(data$value) == totForeigners)

# get iso2c and continent
data$iso2 <- countrycode(data$nationality, "country.name", "iso2c")

data$continent <- countrycode(data$iso2, "iso2c", "continent")
# hack continent for Taiwnan
data[which(data$iso2 == "TW"), 'continent'] <- "Asia"


## Aggregate unfrequent nationalites to "Other countries"
hist(data$value, breaks = 1000)
data[data$value < n.threshold.merge, 'nationality'] %>% unlist(use.name = F) -> countries.sub



# replace unfrequent country by "autres"
data %<>% mutate(nationality = replace(nationality, nationality %in% countries.sub, "Other countries"))

# shape the data for plot, recompute total values
data %<>% group_by(nationality, continent) %>% 
  summarise(value = sum(value, na.rm = T)) %>% ungroup()
stopifnot(sum(data$value) == totForeigners)
data$iso2 <- countrycode(data$nationality, "country.name", "iso2c")

# manual hack for Kosovo
data[which(is.na(data$continent)),]
data[which(data$nationality == "Kosovo"), c('continent', 'iso2')] <- c('Europe', 'KH') 
```

```{r sunburst plot}
lang <- 'FR'

for(lang in 'FR') {#colnames(txt)) {
  cat("\n", lang)
  df <- data
  
  df$country <- countryTranslation(df$iso2, lang)[,-1]
  df[which(df$nationality == "unknown"), 'country'] <- txt['inconnu.country', lang]
  df[which(df$nationality == "Other countries"), 'country'] <- txt["autrespays.country", lang]
  stopifnot(!any(is.na(df$country)))
  
  df$cont <- txt[paste0(df$continent,".continent"), lang]
  df[which(df$nationality == "unknown"), 'cont'] <- txt['inconnu.continent', lang]
  stopifnot(!any(is.na(df$cont)), sum(df$value) == totForeigners)
  
  # create the sequence data
  seqd <- data.frame(
    sequence = paste(df$cont, gsub("(\\-|\\,)", " ", df$country), sep = "-"), 
    freq = df$value
  )
  
  schart <- seqd %>>% sunburst(
    count = T, colors = c(rev(swiTheme::swi_rpal[1:16]),
    rep(swiTheme::swi_rpal[1:19], 8)), legendOrder = "", height = "570px")
  
  # hack
  schart$sizingPolicy$browser$padding <- 0
  schart$sizingPolicy$browser$defaultWidth <- "100%"
  
  saveWidget(schart, file = "swissForeigners_sunburst_tmp.html", selfcontained = FALSE, libdir = "js")
 
  html.grabbed <- grab_widgetHTML("swissForeigners_sunburst_tmp.html")
  html.code <- html.grabbed[['html']]
  istart <- min(grep("script", html.code))
  iend <- max(grep("</head>", html.code))
  header <- html.code[istart:iend-1]
  
  footer <- paste0(
    txt['source',lang], ": ", htmlLink("https://www.pxweb.bfs.admin.ch/Selection.aspx?px_language=fr&px_db=px-x-0103010000_101&px_tableid=px-x-0103010000_101%5Cpx-x-0103010000_101.px&px_type=PX", 
    txt['source.name', lang]),
    " | ",  txt['code', lang], ": ", htmlLink("https://github.com/timelyportfolio/sunburstR", "sunburstR"), " | ",
    htmlLink("http://www.swissinfo.ch", "swissinfo.ch"), " | ",
    htmlLink("https://twitter.com/duc_qn", "@duc_qn")
  )
  out.file <-  paste0("swissForeigners_sunburst_", lang, ".html")
  
  save_html(
    tags$html(
      tags$head(HTML(paste0(header, collapse =""))),
      tags$body(
        h2(HTML(txt['title', lang])), 
        div(class = "descr", HTML(paste0(txt['descr', lang], "<br><br>","<i>", txt['descr2', lang], "</i>"))),
        div(class = "container", HTML(html.code[html.grabbed[['istart']]:html.grabbed[['iend']]])),
        div(id = "cite", HTML(footer)),
        HTML(iframeresizer)
      )), file = out.file, libdir = "js"
  )
  # copy sunburst css
  swi_libWidget_overwrite()
  
    
  # copy my special sunburst with no number abbreviations and no legend
  original <- list.files("js", "sunburst.js", full.names = T, recursive = T)
  file.copy( from = list.files(system.file("extdata", package="swiRcharts"),
    "sunburst_swi.js", full.names = T), to = original, overwrite = T)
  
  # remove legend checkbox
  x <- readLines(out.file)
  y <- gsub('<div class="sunburst-sidebar">         <input type="checkbox" class="sunburst-togglelegend">Legend</input>         <div class="sunburst-legend" style="visibility:hidden;"></div>       </div>', "", x, fixed = T)
  cat(y, file = out.file, sep="\n")
  
  
}
```


### Open the HTML files and remove this code (find & replace):
> <div class="sunburst-sidebar">         <input type="checkbox" class="sunburst-togglelegend">Legend</input>         <div class="sunburst-legend" style="visibility:hidden;"></div>       </div> 

### Repalce sunburst.js in sunburst-binding-0.1: by the edited version. Changes are
> \+ ' / ' + d3.format("1.2s")(totalSize),

Replaced by 

> d3.format("1.2")(d.value),

To change the formrating inside the sunburst

And decrease the transition time to 400 ms


